Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices

### 2022-07-18 paper.tex
[add] initial section on TEE

%-------------------------------------------------------------------------- sub
\section{\uppercase{Discussion}}
\label{sec:discussion}
There are a few places where \mEdhoc{} can be improved,
which we found during this work and communicated to the authors.
%
We discuss them below.
%

>>>>

%-------------------------------------------------------------------------- sec
\section{\uppercase{Trusted Execution Environments}}
\label{sec:TEE}
The main purpose of Trusted Execution Environments (TEE) is to prevent an
adversary with physical access to a device from accessing the computational
circuits.
%
They come in various forms of differing complexity.
%
Some, like ARM Trust Zone and Intel SGX are general purpose execution
environments, which are flexibly programmable.
%
Other, like the Universal Subscriber Identity Modules (USIM) used for
authentication to 3GPP mobile networks, have application specific interfaces for
authentication and key agreement protocols etc.
%
A key aspect is how much of the application is placed in the TEE and how much is
outside.
%

For larger devices that include general purpose processors with Trust Zone or
SGX, entire \mEdhoc{} and \mOscore{} may reside inside the TEE.
%
For constrained IoT devices in the lower end of the scale, a TEE may have to be
implemented using a special-purpose integrated circuit.
%
In the latter case, it may be beneficial to follow a minimalistic approach, and
contain the long-term key and only
the operations that need access to it in the TEE for cost reasons.
%

It may at first appear more secure to implement as much as possible inside the
TEE, but there is a security trade-off: the more code inside the TEE,
the higher the risk of implementation errors in the security-critical code.
%
Because the security critical code runs in the area where the long-term keys
reside, an implementation error here risks leaking information of the key to the
adversary.
%
From this perspective, it may be beneficial to follow the minimalistic approach
even when having access to Trust Zone or SGX.
%

A slightly more secure division of functionality is to also keep the
session key inside
the TEE and extend the interface to accept messages and return the
(en/de)crypted counter part.
%
That is, the interface exposes \mAead{} functions in the interface.
%

%-------------------------------------------------------------------------- sub
\subsection{Practical Aspects and Use Case}
\label{sec:TEE:useCase}
Authentication of a specific IoT device assumes that this devices is the only
entity with access to the long-term key associated with the corresponding
identity.
%
Since IoT devices may be accessible to adversaries, e.g., an insider cloning a
key card, the long-term keys must be appropriately protected.
%
A state of the art approach is to use a Trusted Execution Environment (TEE),
which holds the key and provides an API for operations using the key.
%
This is the approach taken by the
Trust-Zone based \mMuEdhoc{}~\cite{DBLP:conf/codaspy/Hristozov0XFLS21}
implementation for example.
%
Typical operations include signatures using the long-term private key of a
party.
%

%-------------------------------------------------------------------------- sub
\subsection{System and Adversary Model Extensions Supporting TEE}
\label{sec:TEE:advModel}
%
We extend the adversary model of~\cite{Norr21} by allowing the adversary
access to a TEE interface on a party of their choosing before and after the
\mEdhoc{} session.
%
This corresponds to a so-called weak post-compromise security
(PCS)~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
%

We split the \mEdhoc{} functionality as follows.
%
The TEE contains the long-term key and allows the non-TEE parts of the
application to perform the operations on it via its interface.
%
More precisely, parties using the \mSig{} authentication method use a TEE with
an interface which accepts a message and returns the signature of that message
using the party's private long-term key.
%
Parties using the \mStat{} authentication method use a TEE with an interface
accepting $g^x$ and returning $g^x$ raised to the party's private long-term key.
\knote{It should accept any value. Adversary may not stick to elements of the
    form $g^x$. Or we explicitly state that the TEE checks that the received
input lies on the curve (but this may not be sufficient: are the curves cyclic
groups? Need to check).}
%
The non-TEE parts implement \mEdhoc{} using this interface.
%
This interface requires the least functionality from the TEE, reducing the TEE's
complexity and possibly cost of its implementation.
%
The functional split is suitable even for the situation where a
constrained device has implemented only the storage of the long-term key in a
special purpose circuit with minimal processing functionality.
%
Since \mEdhoc{} focuses on constrained IoT devices, it seems appropriate to
cater for this setting.
%

%-------------------------------------------------------------------------- sub
\subsection{Extended Formalism and Security Properties}
\label{sec:TEE:fmAndProps}
Formally, we model the TEE interface by adding two new rewrite rules:
%
\begin{small}
\begin{verbatim}
rule forge_SIG:
   [!LTK_SIG($A, ~ltk), In(xx)] --[TEE($A)]-> [Out(sign(xx, ~ltk))]!>

rule exp_STAT:
   [!LTK_STAT($A, ~ltk), In('g'^~xx)] --[TEE($A)]-> [Out(('g'^~xx)^~ltk)]
\end{verbatim}
\end{small}
%
These rules allow the adversary to obtain terms representing signatures on a
value of their choice to forge signatures (\verb|forge_SIG|), or to obtain terms
representing a curve point of their choice raised to the power of the
long-term key (\verb|exp_STAT|).
%

Because it is a trivial attack when the adversary access these rules with values
from the test session, we must disqualify those rule applications.
%
We do so by creating an action fact \verb|TEE($A)|, where \verb|$A$| is the
identity corresponding to the private key used, and then augmenting the
properties with a condition that no such action fact exists from the start of
the protocol execution and its end.
%
Care must be taken when specifying the start and the end.
%
Specifically, ...
\knote{Need to check this. Should we not allow access to both parties keys
    before and after session? How could we otherwise claim KCI?
}
\knote{Need to check whether auth properties are correct. $\forall$ constructs
    should be like the $\exists$ constructs of PFS. Not only for symmetry, but
    they look incorrect now (should not be $\lor \forall \ldots$).
}

%-------------------------------------------------------------------------- sec
\section{\uppercase{Discussion}}
\label{sec:discussion}
There are a few places where \mEdhoc{} can be improved,
which we found during this work and communicated to the authors.
%
We discuss them below.
%
